---
title: "Creating splist stats master list"
output: html_notebook
---


We'll try to use splist stats to do this

###Algorithm

    masterfile = load(endfile)
    for file in endfile-1 to startfile:
      load(file)
        for each entry in file
          If there's not an entry in masterfile
            entry = handle restart issues (entry)
            add entry to masterfile
            
            
### Raw code for debugging

Here's the master algorithm - some more investigative code follows this

```{r}

source("../R/splist_stats.R")

#arguments
start <- 280000
end <- 1960000
step <- 20000
smfroot <- sprintf("%s/out3/",froot)
#body

ts <- seq(end,start,-step)
msp <- splist_stats(sprintf("%ssplist%d.dat",smfroot,ts[1]))
for( tt in 2:length(ts)){
  message(sprintf("Processing file %d",ts[tt]))  
  nnewreactions <- 0
  nnewparents <- 0
  tsp <- splist_stats(sprintf("%ssplist%d.dat",smfroot,ts[tt]))
  tsp$addentry <- F
  restart <-F
  if(ncol(tsp)==5)restart<-T
  #NB This could be made more efficient but I want to debug it
  for(ee in 1:nrow(tsp)){
    
    entry<-msp[tsp$spp[ee] == msp$spp,]
    if(nrow(entry)>0){
      #message(sprintf("Found %d matches for %s",nrow(entry),tsp$spp[ee]))
      entry<-entry[(tsp$act[ee] == entry$act) & (tsp$pass[ee] == entry$pass),]
      if(nrow(entry)>0){
      }
      else{
        
        #message(sprintf("Found no matches for line %d:  %d,%d,%d"
        #                ,ee,tsp$spp[ee],
        #                tsp$act[ee],tsp$pass[ee]))
        tsp$addentry[ee]<-T
        nnewparents <- nnewparents+1
      }
    }
    else{
      #message(sprintf("%d No match for %s",ee,tsp$spp[ee]))
      tsp$addentry[ee]<-T
      nnewreactions <- nnewreactions+1
    }
  }
  message(sprintf("found %d new reactions and %d new parents",nnewreactions,nnewparents))
}


#return
```

- Going back in time, the only time missing entries appear is when we go over a restart
- The splist that's there at restart has some species in that arent in the subsequent splist. I suspect these are spp that aren't involved in any subsequent reactions
- This means we can speed up processing by only seeking to add new entries when a restart is found. 

##Detecting the last list before restart

```{r}

source("../R/splist_stats.R")
#arguments
start <- 300000
end <-   340000
step <-   20000
smfroot <- sprintf("%s/out3/",froot)
#body

ts <- seq(end,start,-step)
msp <- splist_stats(sprintf("%ssplist%d.dat",smfroot,ts[1]))
restartfound <-F
for( tt in 2:length(ts)){
  message(sprintf("Processing file %d",ts[tt])) 
  fn <- sprintf("%ssplist%d.dat",smfroot,ts[tt])
  tsp <- splist_stats(fn)
  if(restartfound){
    
    #Add the missing entries from the pre-restart file:
    
    mspuniq <- cbind(msp$spp,msp$act,msp$pass,msp$seq)
    tspuniq <- cbind(tsp$spp,tsp$act,tsp$pass,tsp$seq)
    
    hits <- tsp[mspuniq %in% tspuniq,]
    misses <- tsp[!(mspuniq %in% tspuniq),]
    
    
    
    
    
    
    #Handle the missing spp in the restart file:
    
    message(sprintf("File %s is the last before a restart",fn))
    restartfound<-F
    
    missingsp <- unique(rsp[!(rsp$spp %in% tsp$spp),])
    message(sprintf("\t%d missing species at restart:",nrow(missingsp)))
    #for(mm in 1:nrow(missingsp)){
    #  message(sprintf("\t\r%d of %s",tsp$n3[mm],tsp$seq[mm]))
    #}
    missingsp <- unique(missingsp[!(missingsp$spp %in% nsp$spp),])
    message(sprintf("\t%d missing species in subsequent splist",nrow(missingsp)))
    for(mm in 1:nrow(missingsp)){
      message(sprintf("\t\r%d of %s",missingsp$n3[mm],missingsp$seq[mm]))
    }
    
    message(sprintf("\tTotal of %d molecules lost",sum(missingsp$n3)))
  }
  if(ncol(tsp)==5){
    restartfound<-T
    #This is needed to check that all spp are accounted for (they won't be)
    rsp <- tsp
    nsp <- splist_stats(sprintf("%ssplist%d.dat",smfroot,ts[tt-1]))
  }
  
}
```
