---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---


Before going straight for the reaction classes, we might want to look at the distribution of deterministic binding and execution in a conf file. So let's do that now: 


```{r}

  datafile <- system.file("extdata", "out1_880000.conf", package = "Rstringmol")
  data <- rconf_rdata(datafile)

  actset <- unique(data$actseq)
  for(aa in 1:length(actset)){
  
    
    
  }



```










We're going to try and figure out reaction classes based *solely* on the .conf file (and not the splist file)

```{r}
  require(stringr)
  datafile <- system.file("extdata", "out1_880000.conf", package = "Rstringmol")
  data <- rconf_rdata(datafile,summarize = F)

  actset <- unique(data$actseq)
  for(aa in 1:length(actset)){
    passet <- unique(data$passeq[data$actseq == actset[aa]])
    for(pp in 1:length(passet)){
      ty <- reaction_type(actset[aa],passet[pp])
      data$type[data$actseq == actset[aa] & data$passeq == passet[pp]] = ty$rtype
      data$detBind[data$actseq == actset[aa] & data$passeq == passet[pp]] = ty$deterministicBind
      data$detExec[data$actseq == actset[aa] & data$passeq == passet[pp]] = ty$deterministicExec
      
    }
  }
  
  tt <- table(data$type)
  tt <- tt[rtypes]
  
  par(mar=c(10,6,1,0))
  barplot(tt,las=2,col=rcols)
  
```

OK, that seemed to work. Let's now try and make a grid of this: 





```{r fig.height=8, fig.width=12}
ps <- rconf_params(datafile)

plot(NA,xlim=c(0,ps$gridx+50),ylim=c(0,ps$gridy),asp=1)
rect(xleft = 0,xright=ps$gridx,ybottom = 0,ytop=ps$gridy,col="black",lty = "blank")

for(mm in 1:nrow(data)){
 xx <- data$actx[mm]
 yy <- data$acty[mm]
 rc <- rcols[which(rtypes == data$type[mm])]
 rect(xleft = xx,xright = xx+1,ybottom = yy,ytop=yy+1,col=rc,lty = "blank")
 xx <- data$passx[mm]
 yy <- data$passy[mm]
 rect(xleft = xx,xright = xx+1,ybottom = yy,ytop=yy+1,col=rc,lty = "blank")
 legend("topright",pch=15,col = rcols,legend=rtypes)
  
}



```



Set the `froot` variable to the path to the files for these tests:

```{r}
#froot="D:/sjh/stringmol/out3smsp/"
#froot="D:/sjh/stringmol/smsp/1705smsp/out5/"
froot="~/Desktop/paulien/sm250r/out3/"
```


ok, that worked! So let's try it for a whole run...

```{r}

tstep <- 20000 #sometimes its 10000
tt <- tstep
sps <- list()
nsps <- 1
gotdata<-T
while(gotdata){
  fn <- sprintf("%sout1_%d.conf",froot,tt)
  
  if(file.exists(fn)){
  
    #Get the data...
    data <- rconf_rdata(fn,summarize = F)
  
    actset <- unique(data$actseq)
    for(aa in 1:length(actset)){
      passet <- unique(data$passeq[data$actseq == actset[aa]])
      for(pp in 1:length(passet)){
        ty <- reaction_type(actset[aa],passet[pp])
        data$type[data$actseq == actset[aa] & data$passeq == passet[pp]] = ty$rtype
        data$detBind[data$actseq == actset[aa] & data$passeq == passet[pp]] = ty$deterministicBind
        data$detExec[data$actseq == actset[aa] & data$passeq == passet[pp]] = ty$deterministicExec
        
      }
    }
    
    # Now do the plot...
    ps <- rconf_params(fn)
    
    plot(NA,xlim=c(0,ps$gridx+50),ylim=c(0,ps$gridy),asp=1)
    rect(xleft = 0,xright=ps$gridx,ybottom = 0,ytop=ps$gridy,col="black",lty = "blank")
    
    for(mm in 1:nrow(data)){
     xx <- data$actx[mm]
     yy <- data$acty[mm]
     rc <- rcols[which(rtypes == data$type[mm])]
     rect(xleft = xx,xright = xx+1,ybottom = yy,ytop=yy+1,col=rc,lty = "blank")
     xx <- data$passx[mm]
     yy <- data$passy[mm]
     rect(xleft = xx,xright = xx+1,ybottom = yy,ytop=yy+1,col=rc,lty = "blank")
     legend("topright",pch=15,col = rcols,legend=rtypes)
      
    }
  }
  else{
    gotdata<-F
  }
  
  tt <- tt+tstep
}
```

